---
title: "QC"
output:
  html_document:
    df_print: paged
---



```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

```



```{r, libraries}

library(Seurat)
library(cowplot)
library(SCFunctionsV3)
library(dplyr)
library(tidyr)
library(stringr)
library(tidyverse)
library(DoubletFinder)
library(pheatmap)


### Functions
howManyCells <- function(sObj){dim(sObj)[2]}


### Functions
preDoublets <- function(seuratObj){  
    seuratObj <- NormalizeData(seuratObj)
    
    seuratObj <- ScaleData(seuratObj)
    seuratObj <- FindVariableFeatures(seuratObj, 
                                      x.low.cutoff = 0.0125, 
                                      y.cutoff = 0.25, 
                                      do.plot=FALSE)
    seuratObj <- RunPCA(seuratObj, pc.genes = seuratObj@var.genes, pcs.print = 0)
    
    # set.seed(1234)
    # seuratObj <- RunTSNE(seuratObj, dims.use = 1:10, verbose=TRUE)
    return(seuratObj)
}

findDoublets <- function(seuratObj){
    
    ### calculate expected number of doublets
    nExp_poi  <- round(0.15*nrow(seuratObj@meta.data))
    
    ### predict doublets
    seuratObj <- doubletFinder_v3(seuratObj, 
                                  PCs = 1:10, 
                                  pN = 0.25, 
                                  pK = 0.01, 
                                  nExp = nExp_poi, 
                                  reuse.pANN = FALSE, 
                                  sct=FALSE)
    
    
    seuratObj@meta.data <- seuratObj@meta.data %>% 
        rename_at(vars(starts_with("DF.classifications")), 
                  funs(str_replace(., ".*", "DF.classifications"))) %>%
        rename_at(vars(starts_with("pANN")), 
                  funs(str_replace(., ".*", "pANN")))
    
    return(seuratObj) 
    
}



```

```{r}
# after processing
sObj_List <- readRDS("rObjects/sObj_List_all.rds")
sObj_Filtered <- readRDS("rObjects/sObj_Filtered_all.rds")

```

# Samples

```{r}
sampleInfo <- readRDS("rObjects/sampleInfo_all.rds")
sampleInfo[,]

```

The sequencing data was aligned to human reference genome (GRCh38) using CellRanger Version 3.0.1.


```{r, eval = FALSE}
# 1_QC.R
```


### QC

It is important to exclude dead cells and debris from downstream analysis.  This can be done by filtering cells which have low number of reads or genes or a high percentage of mitochondrial reads. The plots below were used to set filtering threshold for the data.

```{r}
nCount_RNA_List <- list()
nFeature_RNA_List <- list()
mt_List <- list()


minCounts <- 1000
minFeatures <- 500
maxMit <- 10


for (i in 1:nrow(sampleInfo)){
  nCount_RNA_List[[i]] <- QCplotHist(sObj_List[[i]], "nCount_RNA", Xmin =  0  , Xmax = 80000,
    NumberOfBins = 20) + xlim(minCounts, 80000) + ggtitle(sampleInfo$fName[i])
  
  nFeature_RNA_List[[i]] <- QCplotHist(sObj_List[[i]], "nFeature_RNA",  Xmax = NULL, Xmin = minFeatures,
    NumberOfBins = 15) + ggtitle(sampleInfo$fName[i])
  
  mt_List[[i]] <- FeatureScatter(object = sObj_List[[i]], 
                 feature1 = "nFeature_RNA", 
                 feature2 = "percent.mt") + 
       geom_hline(yintercept = maxMit) + ggtitle(sampleInfo$fName[i])
   
}
```


## Number of Counts per cell

```{r, fig.height=20, fig.width = 20}
plot_grid(plotlist = nCount_RNA_List, ncol = 1)
```

## Number of Gene per cell

```{r, fig.height=20, fig.width = 20}
plot_grid(plotlist = nFeature_RNA_List, ncol = 1)
```

## Percentage mitochondrial reads

```{r, fig.height=15, fig.width = 20}
plot_grid(plotlist = mt_List, ncol = 2)
```


### Filtering

The following filtering thresholds were applied to the data:    

minimum counts per cell: 1000   
minimum genes per cell: 500   
maximum percentage mitochondria: 10%    

```{r, eval = FALSE}
# 1_QC.R
```




The number of cells for each sample before and after filtering is shown below:

```{r}
sampleInfo
```





