---
title: "AD3 all timepoints, reclustering"
output:
  html_document:
    df_print: paged
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r libaries}
library(dplyr)
library(Seurat)
library("ggplot2")
library(monocle3)
library(patchwork)
library(SeuratWrappers)
library(harmony)
library(ggplotify)
library(tidyverse)

## source markers and plot function
source("../../Retinal_expression_plots.R")

sobjToCds <- function(sObj){
  harmonyLoadings <- sObj@reductions[["harmony"]]@feature.loadings.projected
  sObj <- DietSeurat(sObj, graphs = "umap")
  cds <- as.cell_data_set(sObj)
  cds <- cluster_cells(cds, reduction_method = "UMAP", k = 15)
  cds <- learn_graph(cds, use_partition = FALSE)
  ## Add gene names into CDS
  rowData(cds)$gene_name <- rownames(cds)
  rowData(cds)$gene_short_name <- rowData(cds)$gene_name
  ## estimate size factors
  cds <- estimate_size_factors(
  cds,
  round_exprs = TRUE,
  method = c("mean-geometric-mean-total", "mean-geometric-mean-log-total")
)
  # add harmony reduction
  cds@preprocess_aux$gene_loadings <- harmonyLoadings
  
  
return(cds)
}


seuratObj <- readRDS("rObjects/seuratObj_integrated.rds")




```


# Reclusting {.tabset .tabset-fade .tabset-pills}

## Cluster 13/1/22

The integrated data (from 5_inegtation_AD3.Rmd) containing all AD3 samples was annotated using the powerpoint Ad3 organoids bh+ll sent on 2/11/21 (Slide 28)

### Annotations

```{r}
outName <- "all_samples_clusters_removed"
annotations <- read.csv(paste0("annotations/", outName, "_annotation"))
annotations


clustersToRemove <- annotations$Cluster[grepl("removed_", annotations$Cell_Type)]
seuratObj_subset <- seuratObj[,!seuratObj@active.ident %in% clustersToRemove]
annotations$Cluster <- factor(annotations$Cluster)
seuratObj_subset$annotation <- annotations$Cell_Type[match(seuratObj_subset@active.ident, annotations$Cluster)]
```

### Annotated clusters

```{r}
DimPlot(seuratObj_subset, group.by = "annotation")
```


The data was then reprocessed using harmony batch correction and cell cycle regression, and re-clustered. There were 36 clusters. The cells seem to still be grouped by the cell type annotation used previously. Marker genes were found for each of the new clusters.
(Linda and Birthe Analysis on slide 33)

```{r, eval = FALSE}
## cell cycle genes
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes



seuratObj_reclustered <- seuratObj_subset %>% 
  CellCycleScoring(s.features = s.genes, 
                   g2m.features = g2m.genes, 
                   set.ident = FALSE) %>% 
  Seurat::NormalizeData(verbose = FALSE) %>%
  FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>% 
  ScaleData(verbose = FALSE, 
            vars.to.regress = c("nCount_RNA", "nFeature_RNA", "percent.mt")) %>% 
  RunPCA(pc.genes = seuratObj_ds@var.genes, npcs = 20, verbose = FALSE) %>%
  RunHarmony(c("batch", "seq_batch", "orig.ident"), plot_convergence = FALSE) %>%
  RunUMAP(reduction = "harmony", dims = 1:10) %>%
  FindNeighbors(reduction = "harmony", dims = 1:10) %>% 
  FindClusters(resolution = c(0.2,1,2.2))


## Find Markers
markers <- FindAllMarkers(seuratObj_reclustered, only.pos = TRUE, logfc.threshold = 0.5 )
markers <- markers %>% arrange(cluster, desc(avg_log2FC))

fName <- paste0("csvFiles/markers_annotation_", outName, ".csv")
write.csv(markers, fName)
fName <- gsub("csvFiles", "rObjects", fName) %>% gsub(".csv", ".rds", .)

saveRDS(seuratObj_reclustered, paste0("rObjects/seurat_", outName, ".rds"))
```


### New Clusters

```{r, fig.width=20, fig.height = 15}
seuratObj_reclustered <- readRDS(paste0("rObjects/seurat_", outName, ".rds"))
p1 <- DimPlot(seuratObj_reclustered, label = TRUE) + ggtitle("New clusters")
p2 <- DimPlot(seuratObj_reclustered, group.by = "annotation", label = TRUE, repel = TRUE) + ggtitle("Previous annotation")
p3 <- DimPlot(seuratObj_reclustered, group.by = "day", label = TRUE) + ggtitle("Timepoint")

p1 
p2 
p3

```





## Clusters 17/1/22 (with 20)

### Clusters Removed

17/1/22 The data was the annotated by Linda and Birthe (slide 33) and reclustered.

```{r}
annotations <- read.csv("annotations/all_samples_17_1_2022")
annotations
```

```{r,fig.width=15, fig.height = 15}

clustersToRemove <- annotations$Cluster[grepl("removed_", annotations$Cell_Type)]
DimPlot(seuratObj_reclustered[,!seuratObj_reclustered@active.ident %in% clustersToRemove], label = TRUE, repel = TRUE)
```


```{r, eval = FALSE}
# Reclustering_17_1_22.R
```

### New Clusters

```{r, fig.width=15, fig.height = 15}
outName <- "17_1_22"
seuratObj_reclustered_with <- readRDS(paste0("rObjects/seurat_", outName, ".rds"))

p1 <- DimPlot(seuratObj_reclustered_with, label = TRUE) + ggtitle("New clusters")
p2 <- DimPlot(seuratObj_reclustered_with, group.by = "annotation", label = TRUE, repel = TRUE) + ggtitle("Previous annotation")
p3 <- DimPlot(seuratObj_reclustered_with, group.by = "day", label = TRUE) + ggtitle("Timepoint")

p1 
p2 
p3

```


## Pluripotent Stem Cells

Add modules was used to calculate average expression for "L1TD1",
"POU5F1",
"DNMT3B",
"TERF1",
"DPPA4",
"LIN28A"

Cells with an expression greater than 1 were then removed from the analysis (44889/48585)



```{r, eval = FALSE}
pluriGenes <- list(c("L1TD1",
"POU5F1",
"DNMT3B",
"TERF1",
"DPPA4",
"LIN28A"))



seuratObj_reclustered_with <- AddModuleScore(seuratObj_reclustered_with, features = pluriGenes, name = "Pluripotent")
outName <- "17_1_22"
saveRDS(seuratObj_reclustered_with,
            paste0("rObjects/seurat_", outName, ".rds"))









```

```{r}
FeaturePlot(seuratObj_reclustered_with, "Pluripotent1", min.cutoff = 0, max.cutoff = 1)
```

The data was then reprocessed using harmony batch correction and cell cycle regression, and re-clustered.  Marker genes were found for each of the new clusters.


```{r, eval = FALSE}

seuratObj_pluripotent_removed <- seuratObj_reclustered_with[,seuratObj_reclustered_with$Pluripotent1 < 1]

## cell cycle genes
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes



seuratObj_pluripotent_removed <- seuratObj_pluripotent_removed %>% 
  CellCycleScoring(s.features = s.genes, 
                   g2m.features = g2m.genes, 
                   set.ident = FALSE) %>% 
  Seurat::NormalizeData(verbose = FALSE) %>%
  FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>% 
  ScaleData(verbose = FALSE, 
            vars.to.regress = c("nCount_RNA", "nFeature_RNA", "percent.mt")) %>% 
  RunPCA(pc.genes = seuratObj_ds@var.genes, npcs = 20, verbose = FALSE) %>%
  RunHarmony(c("batch", "seq_batch", "orig.ident"), plot_convergence = FALSE) %>%
  RunUMAP(reduction = "harmony", dims = 1:10) %>%
  FindNeighbors(reduction = "harmony", dims = 1:10) %>% 
  FindClusters(resolution = c(0.2,1,2.2))

outName <- "clusters_and_pluripotent_removed"
saveRDS(seuratObj_pluripotent_removed, paste0("rObjects/seurat_", outName, ".rds"))
## Find Markers
markers <- FindAllMarkers(seuratObj_pluripotent_removed, only.pos = TRUE, logfc.threshold = 0.5 )
markers <- markers %>% arrange(cluster, desc(avg_log2FC))

fName <- paste0("csvFiles/markers_annotation_", outName, ".csv")
write.csv(markers, fName)
fName <- gsub("csvFiles", "rObjects", fName) %>% gsub(".csv", ".rds", .)


```

### New Clusters

```{r, fig.width=20, fig.height = 15}
outName <- "clusters_and_pluripotent_removed"

seuratObj_pluripotent_removed <- readRDS(paste0("rObjects/seurat_", outName, ".rds"))
p1 <- DimPlot(seuratObj_pluripotent_removed, label = TRUE) + ggtitle("New clusters")
p2 <- DimPlot(seuratObj_pluripotent_removed, group.by = "annotation", label = TRUE, repel = TRUE) + ggtitle("Previous annotation")
p3 <- DimPlot(seuratObj_pluripotent_removed, group.by = "day", label = TRUE) + ggtitle("Timepoint")

p1 
p2 
p3

```