---
title: "Monocle RPCs T1 HCs ACs RGCs - organoids"
output:
  html_document:
    df_print: paged
---


```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

```


```{r, libraries}
library(tidyverse)
library(Signac)
library(Seurat)
library(JASPAR2020)
library(GenomeInfoDb)
library(EnsDb.Hsapiens.v86)
library(TFBSTools)
library(monocle3)
library(SeuratWrappers)
library(patchwork)

pfm <- getMatrixSet(
  x = JASPAR2020,
  opts = list(collection = "CORE", tax_group = 'vertebrates', all_versions = FALSE)
)

# motifs <- sapply(pfm, function(x){x@name})
# motif_df <- data.frame(motif = names(motifs),
#            name = unname(motifs)
#            )
# 
# 
# retina_markers <- readRDS("../markers/Retina_Markers.rds")
# #retina_markers_df <-
# 
# 
# motif_df$motif[match(retina_markers[[2]], motif_df$name)]
motifs <- readRDS("rObjects/shared_enriched_motifs_cellTypes.rds")
sObj <- readRDS("rObjects/shared_sObj_merge_retina_with45_60.rds")
# DimPlot(integrated, label = TRUE)


motif_expr <- sObj@assays$chromvar@data
motif_expr <- motif_expr[rowSums(motif_expr) > 0 ,]
sObj$nMotif <- motif_expr %>% colSums()

sObj$CellType <- sObj$CellType %>% as.character()
## remove space
sObj$CellType[grepl("MG", sObj$CellType)] <- "MG"
cellorder <- c("pCM", "early RPCs","late RPCs", "T1","T2","T3","RGCs", "HCs + ACs", "cones", "rods", "BPs", "microglia", "MG")

sObj$CellType <- factor(sObj$CellType, levels = cellorder)

sObj <- SetIdent(sObj, value = "CellType")
```


## Annotate cells

The cells were annotated using the annotations on slide 155 with new AC definitions:

```{r}
DimPlot(sObj, group.by = "CellType", label = TRUE)


```



## RPCs T1 HCs ACs RGCs trajectory

```{r, eval = FALSE}

## subset_for_pseudotime.R


```

```{r}
##variables to change
pseudotime_order <- c("pCM", "early RPCs", "late RPCs", "T1", "RGCs", "HCs + ACs")
start_Node_ATAC <- "Y_96"
atacName <- "pCM_RPC_T1_HCs_ACs_RGCs"
atacTitle <- "pCM - RPC - T1 - RGCs - HCs - ACs"
sObj_ATAC_subset <- readRDS(paste0("rObjects/seurat_PT_",atacName, ".rds"))

##create out dir
dir.create(file.path("highRes", atacName), showWarnings = FALSE)

## create cds object
cds <- as.cell_data_set(sObj_ATAC_subset)
cds <- cluster_cells(cds = cds, reduction_method = "UMAP")
cds <- learn_graph(cds, use_partition = FALSE, close_loop = FALSE)
plot_cells(cds, show_trajectory_graph = TRUE, label_principal_points = TRUE)

cds <- order_cells(cds, reduction_method = "UMAP", root_pr_nodes = start_Node_ATAC)
plot_cells(cds, show_trajectory_graph = TRUE, label_principal_points = TRUE, color_cells_by = "pseudotime")


### add pseudotime back to seurat object
sObj <- AddMetaData(
  object = sObj,
  metadata = abs(cds@principal_graph_aux@listData$UMAP$pseudotime),
  col.name = "pseudotime"
)

```


The RPCs, T1, RGCs cell types were selected and Monocle 3 pseudotime analysis performe starting from `r start_Node_ATAC`.  


```{r}
p1 <- plot_cells(
  cds = cds,
  show_trajectory_graph = TRUE,
  color_cells_by = "CellType", label_principal_points = TRUE, cell_stroke = .5
) 





p1 <- p1
p1$layers[[1]] <- NULL
p1
```



```{r}
p0 <-  plot_cells(
  cds = cds,
  show_trajectory_graph = TRUE,
  color_cells_by = "pseudotime", label_principal_points = TRUE,  cell_stroke = .5
) 

p0$layers[[1]] <- NULL
p0

```

The pseudotime estimated are shown on the original umap below:


```{r}
DimPlot(sObj, group.by = "CellType", label = TRUE)

FeaturePlot(sObj, c("pseudotime"), pt.size = 0.1) & scale_color_viridis_c()

```



The correlation between pseudotime between motifs enrichment was calculated:

```{r}
cellTypes <- names(motifs)
pseudotime <-sObj$pseudotime[!is.na(sObj$pseudotime)]
cor_row <- function(x){cor(pseudotime, x) }


addCorrelation <- function(cellType){
motif_mat <- sObj@assays$chromvar@data[motifs[[cellType]]$motif,!is.na(sObj$pseudotime)]
motifs[[cellType]]$corr_pseudotime <- apply(motif_mat, 1, cor_row)
return(motifs[[cellType]])
}


motifs <- lapply(names(motifs), addCorrelation)
names(motifs) <- cellTypes

```


## Top Motif plots {.tabset .tabset-pills}

The motif lists were filtered to include only motifs in at least 30% of the cells that there were enriched for with a log fold change of over 2. The top 5 absolute correlation scores for each cell type then selected


```{r}
motifs_filtered <- lapply(motifs, function(x) {x[x$fold.enrichment > 2 & x$percent.observed > 30, ]})



findTopMotifs <- function(x){x %>% arrange(desc(abs(corr_pseudotime))) %>% head(n = 5)}

topMotifs <- lapply(motifs_filtered, findTopMotifs)

# 
# 
# no_motifs <- motifs_filtered %>% sapply(nrow) %>% data.frame() 
# colnames(no_motifs) <- "number of motifs"
# 
# no_motifs

```

```{r}
MotifFeaturePlot <- function(motif){
  motif_name <- motifs[[1]]$motif.name[motifs[[1]]$motif == motif]
  p1 <- FeaturePlot(sObj, motif, min.cutoff = 0, cols = c("lightgrey", "red"), order = TRUE, pt.size = 1 ) + ggtitle( paste(motif_name, motif))
  p2 <- MotifPlot(sObj, motif )
 layout <- '
A#
AB
'
p3 <- p1 + p2   +
  plot_layout(design = layout, widths = c(2,1))
  return(p3)
}



DefaultAssay(sObj) <- "peaks"
```



```{r}
motif_names <- sObj@assays$peaks@motifs@motif.names

pCM <- topMotifs[["pCM"]]$motif
earlyRPCs <- topMotifs[["early RPCs"]]$motif
lateRPCs <- topMotifs[["late RPCs"]]$motif
T1 <- topMotifs[["T1"]]$motif
HCs_ACs <- c(topMotifs[["HCs + ACs"]]$motif)
RGCs <- c(topMotifs[["RGCs"]]$motif)



all <- data.frame(motif = unique(c(pCM,earlyRPCs,lateRPCs,T1,HCs_ACs,RGCs)))

all$pCM <- all$motif %in% pCM
all$earlyRPCs <- all$motif %in% earlyRPCs
all$lateRPCs <- all$motif %in% lateRPCs
all$T1 <- all$motif %in% T1
all$HCs_ACs <- all$motif %in% HCs_ACs
all$RGCs <- all$motif %in% RGCs


cell_types <- colnames(all)[-1]
getCluster <- function(x){
y <- paste0(cell_types[x], collapse = "_")
return(y)}

all$cluster <- apply(all[,cell_types], 1, getCluster)

all$motif_names <- motif_names[all$motif] %>%unlist() %>% unname()
```



## pCM 

```{r}
cluster <- "pCM"
motifs_to_plot <- all$motif[all$cluster == cluster]

lapply(motifs_to_plot , MotifFeaturePlot)

# motifs_names_to_plot <- all$motif_names[all$cluster == cluster] %>% intersect(footprints)
# if(length(motifs_names_to_plot) > 0){
# PlotFootprint(sObj, features = motifs_names_to_plot)}
```



## early RPCs 

```{r}
cluster <- "earlyRPCs"
motifs_to_plot <- all$motif[all$cluster == cluster]

lapply(motifs_to_plot , MotifFeaturePlot)

# motifs_names_to_plot <- all$motif_names[all$cluster == cluster] %>% intersect(footprints)
# if(length(motifs_names_to_plot) > 0){
# PlotFootprint(sObj, features = motifs_names_to_plot)}
```

## late RPCs 

```{r}
cluster <- "lateRPCs"
motifs_to_plot <- all$motif[all$cluster == cluster]

lapply(motifs_to_plot , MotifFeaturePlot)

# motifs_names_to_plot <- all$motif_names[all$cluster == cluster] %>% intersect(footprints)
# if(length(motifs_names_to_plot) > 0){
# PlotFootprint(sObj, features = motifs_names_to_plot)}
```

## T1

```{r}
cluster <- "T1"
motifs_to_plot <- all$motif[all$cluster == cluster]

lapply(motifs_to_plot , MotifFeaturePlot)

# motifs_names_to_plot <- all$motif_names[all$cluster == cluster] %>% intersect(footprints)
# if(length(motifs_names_to_plot) > 0){
# PlotFootprint(sObj, features = motifs_names_to_plot)}
```

#lateRPCs_RGCs

```{r}
cluster <- "lateRPCs_RGCs"
motifs_to_plot <- all$motif[all$cluster == cluster]

lapply(motifs_to_plot , MotifFeaturePlot)

# motifs_names_to_plot <- all$motif_names[all$cluster == cluster] %>% intersect(footprints)
# if(length(motifs_names_to_plot) > 0){
# PlotFootprint(sObj, features = motifs_names_to_plot)}
```



## RGCs

```{r}
cluster <- "RGCs"
motifs_to_plot <- all$motif[all$cluster == cluster]

lapply(motifs_to_plot , MotifFeaturePlot)

# motifs_names_to_plot <- all$motif_names[all$cluster == cluster] %>% intersect(footprints)
# if(length(motifs_names_to_plot) > 0){
# PlotFootprint(sObj, features = motifs_names_to_plot)}
```



## HCs_ACs 

```{r}
cluster <- "HCs_ACs"
motifs_to_plot <- all$motif[all$cluster == cluster]

lapply(motifs_to_plot , MotifFeaturePlot)

# motifs_names_to_plot <- all$motif_names[all$cluster == cluster] %>% intersect(footprints)
# if(length(motifs_names_to_plot) > 0){
# PlotFootprint(sObj, features = motifs_names_to_plot)}
```
