---
title: "RPCs_T1_T2_T3_final - to use"
output:
  html_document:
    df_print: paged
---
```{r}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r libaries}
library(dplyr)
library(Seurat)
library("ggplot2")
library(monocle3)
library(patchwork)
library(SeuratWrappers)
library(harmony)
library(ggplotify)
library(PNWColors)
library(pheatmap)
library(viridis)


seuratObj <- readRDS("../rObjects/seurat_clusters_and_pluripotent_removed.rds")
subsets <- readRDS("../rObjects/pseudotime_subsets.rds")


outName <- "RPCs_T1_T2_T3_October"
seuratObj_subset_part1 <- readRDS(paste0("../rObjects/seurat_subset_part1", outName, ".rds"))

cds_selected <- readRDS(paste0("../rObjects/cds_", outName, ".rds"))

# seuratObj_subset <- readRDS(paste0("rObjects/seurat_subset_part2", outName2, ".rds"))
# cds_subset_cc <- readRDS(paste0("rObjects/cds_", outName2, ".rds"))
subsets <- readRDS("../rObjects/pseudotime_subsets.rds")
#startingNode <- read.csv("StartNode/startingNodes", row.names = 1)[outName,]

clusters_to_remove <- readRDS("../rObjects/clutsters_to_remove.rds")[[outName]]

seuratObj_selected <- readRDS(paste0("../rObjects/"," seurat_subset_clusters_removed", outName, ".rds"))
colours <-  read.csv("/data/rachel/Linda_Lako/Retina/HighResFigures/colours_birthe.csv")

```

```{r}
## functions

sobjToCds <- function(sObj){
  harmonyLoadings <- sObj@reductions[["harmony"]]@feature.loadings.projected
  sObj <- DietSeurat(sObj, graphs = "umap")
  cds <- as.cell_data_set(sObj)
  cds <- cluster_cells(cds, reduction_method = "UMAP", k = 15)
  cds <- learn_graph(cds, use_partition = FALSE)
  ## Add gene names into CDS
  rowData(cds)$gene_name <- rownames(cds)
  rowData(cds)$gene_short_name <- rowData(cds)$gene_name
  ## estimate size factors
  cds <- estimate_size_factors(
  cds,
  round_exprs = TRUE,
  method = c("mean-geometric-mean-total", "mean-geometric-mean-log-total")
)
  # add harmony reduction
  cds@preprocess_aux$gene_loadings <- harmonyLoadings
  
  
return(cds)
}


pt_heatmap <- function(seuratObj_subset, cds_subset_cc, scenario) {
  seuratObj_subset <- SetIdent(seuratObj_subset, value = "CellType")
  cellTypes <- seuratObj_subset$CellType %>% unique()


cellTypeOrder <- read.csv("../annotations/cellTypeOrder.csv", 
                          header = FALSE, 
                          stringsAsFactors = FALSE)$V1 

clusterOrder <- unique(cellTypeOrder[cellTypeOrder %in% cellTypes])
seuratObj_subset$CellType2 <- seuratObj_subset$CellType %>% factor(levels = clusterOrder)


seuratObj_subset$pseudotime <- pseudotime(cds_subset_cc)
cellOrder <- order(seuratObj_subset$CellType2, rev(seuratObj_subset$pseudotime))

## colours for legend
pal <- pnw_palette("Bay",length(clusterOrder),type="continuous")[1:length(clusterOrder)]
names(pal) <- clusterOrder
annotation_colors <- list(CellType = pal)


fName <- paste0("celltype_markers/", outName, "scenario", "_", scenario, "_celltype_markers.csv")
markers <- read.csv(fName, row.names = 1)
markers$cluster <- markers$cluster %>%  factor(levels = clusterOrder )




top10 <- markers %>% 
  filter(avg_log2FC > 0) %>% 
  arrange(cluster) %>% 
  group_by(cluster) %>% 
  top_n(n = 10, wt = avg_log2FC) 

#top10 <- top10[c(-1,-2,-3),]

genesUse <- top10 %>% pull(gene) %>% unique()


md <- seuratObj_subset[[]][cellOrder,c("CellType", "pseudotime", "day")]
breaksList = seq(0, 3, by = .2)


exp <- seuratObj_subset@assays$RNA@data[genesUse, cellOrder]

p_heatmap1 <- as.ggplot(pheatmap(exp, 
         cluster_rows = FALSE, 
         cluster_cols = FALSE, 
         show_colnames = FALSE, 
         annotation_col = md, 
         color = viridis(15), 
         annotation_colors = annotation_colors,
         scale = "row", 
         breaks = breaksList
          ))


}

```

# {.tabset .tabset-pills}

## Clusters used


```{r}
clustersUse <- subsets[[outName]]

## subset_trajectory.R
```

Cells from  `r paste0(clustersUse, collapse = ",")` were selected. The cells selected are shown below:

```{r}
seuratObj$cellSelected <- NA

anno <- read.csv("../../../AD3_highRes/annotation/all_anno_October_2022.csv")
seuratObj$anno_oct <- anno$Annotation[match(seuratObj@active.ident, anno$Cluster)]

seuratObj$cellSelected[colnames(seuratObj_subset_part1)] <- seuratObj$anno_oct[colnames(seuratObj_subset_part1)]
seuratObj_subset_part1$annotation <- seuratObj$anno_oct[colnames(seuratObj_subset_part1)]



DimPlot(seuratObj, group.by = "cellSelected")
```


### Clustering and annotation

Cell cycle regression and harmony batch correction performed. The new clusters are shown below:  


```{r, fig.width=25, fig.height = 15}

p1 <- DimPlot(seuratObj_subset_part1, group.by = "RNA_snn_res.1", label = TRUE,  label.size =  5, repel = TRUE)  + ggtitle("new clusters, res.1")

# seuratObj_subset_part1$cellSelected <- NA
# seuratObj_subset_part1$cellSelected[colnames(seuratObj_subset)] <- seuratObj_subset_part1$RNA_snn_res.1[colnames(seuratObj_subset)]

p2 <- DimPlot(seuratObj_subset_part1, group.by = "annotation", label = TRUE, label.size = 5,  repel = TRUE)


p1  + p2 

```


### Further clustering

Clusters `r clusters_to_remove ` were then removed  and the data reclustered:
```{r, fig.width=25, fig.height = 15}

p1 <- DimPlot(seuratObj_selected, group.by = "RNA_snn_res.1", label = TRUE,  label.size =  5, repel = TRUE)  + ggtitle("new clusters, res.1")

p1b  <- DimPlot(seuratObj_selected, group.by = "RNA_snn_res.0.8", label = TRUE,  label.size =  5, repel = TRUE)  + ggtitle("new clusters, res.0.8")

p1c  <- DimPlot(seuratObj_selected, group.by = "RNA_snn_res.0.6", label = TRUE,  label.size =  5, repel = TRUE)  + ggtitle("new clusters, res.0.6")




p1  + p1b + p1c + p2 

```






## Clustered slected
```{r}
clustersSelected <- c(0:8)
startingNode <- "Y_216"
```

Join 8 with cluster 4 (T1),  cluster 1 included. Start node from early RPC CLUSTER 3. Node `r startingNode` ,was used as the starting point for the trajectory. 

### Scenario annotation

```{r, , fig.width= 20}
seuratObj_subset <- seuratObj_selected[, seuratObj_selected$RNA_snn_res.0.6 %in% clustersSelected]
cds_subset_cc <- cds_selected[, colnames(seuratObj_subset)]

anno_new <- read.csv(paste0("annotation/", outName, "_final.csv"), stringsAsFactors = FALSE)
seuratObj_subset$CellType <- anno_new$Annotation[match(seuratObj_subset$RNA_snn_res.0.6, anno_new$Cluster)]
cds_subset_cc$CellType <- seuratObj_subset$CellType

p1 <- DimPlot(seuratObj_subset, group.by = "CellType")
p2 <- plot_cells(cds_subset_cc, 
           label_principal_points = TRUE, 
           color_cells_by = "CellType"
           )

p2 + p1
```

 
```{r}
# convert_seurat_to_cds.R 
```

```{r}
cds_subset_cc <- order_cells(cds_subset_cc, root_pr_nodes = startingNode)
```


### Pseudotime Ordering

The pseudotime ordering is shown below.

```{r, fig.width = 8, fig.height=8}
plot_cells(cds_subset_cc, 
           label_principal_points = FALSE, 
           color_cells_by = "pseudotime", show_trajectory_graph = FALSE
           )
```

### Markers and heatmap for each cluster

Markers were then found for each of the cell types. 

```{r, eval = FALSE}
findmarkers.R
```

```{r, fig.width= 20,fig.height=20}
p1 <- pt_heatmap(seuratObj_subset, cds_subset_cc, scenario = "final" ) 

```



## High Res Plots


## Selected cells 
```{r}


seuratObj$cellSelected[seuratObj$cellSelected %>% grepl("RPCs", .)] <- "RPCs"
cells_included <- seuratObj$cellSelected %>% factor() %>% levels()

colours_use <- colours$Colour[match(cells_included, colours$Annotation)]

p1 <- DimPlot(seuratObj, group.by = "cellSelected") + scale_colour_manual(values = colours_use)



fileName <- "RPCs_T1_T2_T3_selected"
ggsave(paste0("highRes/", fileName, "_umap.tiff"), plot = p1,
         width = 8,
         height = 5,)

```


### UMAP

```{r}
scAnno <- colours$Colour[match(levels(factor(seuratObj_subset$CellType)), colours$Annotation)]

p1 <- DimPlot(seuratObj_subset, label = FALSE, label.size = 5, pt.size = 1, group.by = "CellType") + 
  theme_classic() + 
  scale_colour_manual(values = scAnno)

ggsave(filename = "highRes/RPCs_T1_T2_T3_umap_scenario5.tiff", plot = p1)
```


### heatmap

```{r}
p1 <- pt_heatmap(seuratObj_subset, cds_subset_cc, scenario = "final" ) 
ggsave(paste0("highRes/heatmaps_", outName, ".tiff"), plot = p1, width = 12, height = 10)

```


### Pseudotime
```{r}
p2 <- plot_cells(cds_subset_cc, 
           label_principal_points = FALSE, 
           color_cells_by = "pseudotime", show_trajectory_graph = FALSE
           )

ggsave(filename = "highRes/RPCs_T1_T2_T3_pseudotime_colours_scenario5.tiff", plot = p2)
```
